Viscous Transport in Eroding Porous Media

### 2019-07-16 docs/transport/erosion_transport.tex
working on early parts of paper

Since erosion alters the geometry, it also affects coarse-grained
variables.  In this paper, we examine two particular variables that can
be used to understand transport in porous media---the tortuosity and the
anomalous diffusion rates.   The tortuosity quantifies the twisting of
particle trajectories in a medium, and in laminar flow, the tortuosity
is one.  While a porous media often has a large tortuosity, as erosion
creates preferential channels, large numbers of particles become more
laminar, and we expect the tortuosity to decrease.  Anomalous diffusion
is often used to describe transport in a porous media, and it quantifies
the dispersion of particles in the flow.  Again, as the geometry erodes,
particles behave more as a ballistic particle, and we expect less
dispersion.  In this paper, we compute both the tortuosity and the
anomalous diffusion rates of several geometries that form via erosion.

\todo[inline]{gap size growth as a function of time}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Boundary Integral Equation Formulation}
\label{s:formulation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We start by defining the main variables used to model erosion.  We
consider flows inside a confined geometry $\Omega$ that contains $M$
eroding bodies with boundary $\gamma_\ell$.  The boundary of the
bounding geometry is $\Gamma$, and the fluid domain is bounded by $\bd
\Omega = \Gamma \cup \gamma_1 \cup \cdots \cup \gamma_M$.  On each
eroding body, $\nn$ is the normal vector pointing into the body, and
$\ss$ is the unit tangent vector pointing in the counterclockwise
direction.  The normal vector on $\Gamma$ is the outward unit normal.
Neglecting inertial forces, the governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{\em conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{\em conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{\em no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{\em outer wall
velocity} \\
\Vn = \CE \, \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
&&\mbox{\em erosion model}.
\end{split}
\end{equation}
Here, $\uu$ is the fluid velocity, $p$ is the pressure, $\UU$ is a
prescribed Hagen-Poiseuille velocity field, and
\begin{align}
  \tau = -(\nabla \uu + \nabla \uu^T) \nn \cdot \ss
  \label{eqn:shearStress}
\end{align}
is the shear stress on the eroding bodies.  Continuing with our previous
work~\cite{qua-moo2018}, we use a double-layer potential formulation for
the velocity $\uu$.  The double-layer potential is
\begin{align}
  \uu(\xx) = \DDD[\eeta](\xx) = 
    \int_{\bd\Omega} D(\xx,\yy) \eeta(\yy)\, ds_\yy = 
  \frac{1}{\pi}\int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
  \label{eqn:stokesDLP}
\end{align}
where $D$ is the kernel of the integral operator, $\rr = \xx - \yy$,
$\rho = \|\rr\|$, $\nn$ is the unit outward normal at $\yy$, and $\eeta$
is an unknown density function.  The Stokes double-layer potential
itself can not represent all solutions of the incompressible Stokes
equations.  Therefore, we complete the integral equation formulation
following the work of Power and Miranda~\cite{pow-mir1987}.  The
completion depends on the Stokeslets and rotlets
\begin{align}
  S[\llambda_\ell,\cc_\ell](\xx) = \frac{1}{4\pi} \left( 
    -\log \rho \II + \frac{\rr \otimes \rr}{\rho^2} \right), \quad
  R[\xi_\ell,\cc_\ell](\xx) = \frac{\rr^\perp}{\rho^2} \xi_\ell,
\end{align}
where $\rr = \xx - \cc_\ell$, $\cc_\ell$ is a point inside the
$\ell^{th}$ eroding body, $\rr^\perp = (r_2,-r_1)$, and $\rho =
\|\rr\|$.  Then, the solution of the Stokes equation is
\begin{align}
  \uu(\xx) = \DDD[\eeta](\xx) + 
    \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx) + 
    \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx),
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DDD[\eeta](\xx) + 
    \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx) + 
    \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx) +
    \NN_0(\eeta](\xx), \quad \xx \in \bd\Omega, \\
  \llambda_\ell &= \frac{1}{2\pi} \int_{\gamma_\ell} 
    \eeta(\yy)\, ds_\yy \quad \ell = 1,\ldots,M, \\
  \xi_\ell &= \frac{1}{2\pi} \int_{\gamma_\ell}
    (\yy - \cc_\ell)^\perp \cdot \eeta(\yy)\, ds_\yy 
    \quad \ell = 1,\ldots,M,
\end{align}
where $\ff$ is the prescribed Dirichlet boundary condition of the
velocity on the outer wall and eroding bodies.  Then, once the unknown
density function is computed, the shear stress is found by first
computing the deformation tensors of the double-layer potential,
Stokeslets, and rotlets, and then forming the shear stress as defined in
equation~\eqref{eqn:shearStress}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{s:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Using the trapezoid rule to discretize the double-layer
potential~\eqref{eqn:stokesDLP}, and to compute its deformation tensor,
results in large quadrature errors if $\xx$ is close to $\bd\Omega$.  We
apply a quadrature method that is constructed in terms of holomorphic
functions, so we must first write all the layer potentials in terms of
Cauchy integrals and their derivatives.  The first step is to write the
Stokes double-layer potential as the sum of a Laplacian double-layer
potential
\begin{align}
  \DD[\eta](\xx) = \frac{1}{2\pi} \int_{\bd\Omega}
    \frac{\rr \cdot \nn}{\rho^2} \eta(\yy) ds_\yy,
\end{align}
and its gradient.  One such decomposition is
\begin{equation}
  \label{eqn:Stokes2Laplace}
  \begin{aligned}
    \DDD[\eeta](\xx) &= 
      \frac{1}{2\pi} \int_{\bd\Omega} 
        \frac{\nn}{\rho^2} (\rr \cdot \eeta) \, ds_\yy + 
      \frac{1}{2\pi} \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2} (\yy \cdot \eeta) \, ds_\yy \\
      &- \frac{1}{2\pi} x_1 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_1(\yy) \, ds_\yy -
      \frac{1}{2\pi} x_2 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_2(\yy) \, ds_\yy.
  \end{aligned}
\end{equation}
Next, since the Laplacian double-layer potential can be written as the
Cauchy integral
\begin{align}
  \DD[\eta](\xx) = \Re (v(x)) \quad \text{where} \quad
  v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy,
  \label{eqn:laplaceComplex}
\end{align}
where $\xx,\yy \in \RR^2$ are identified with the points $x,y \in \CC$,
the Stokes double-layer potential can be written in terms of Cauchy
integrals and its first derivative.

We also require the derivative of
the Stokes double-layer potential which requires the second derivative
of Cauchy integrals.

We will be using quadrature formulae to compute $v$ and its derivatives
that rely on the Cauchy integral theorem.  Since $v$ is holomorphic, its
value only depends on its boundary data.  The boundary data satisfies
the Sokhotski-Plemelj jump relation
\begin{align}
  \label{eqn:SPrelation}
  v(x) = -\frac{1}{2} \eta(x) + \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x-y}\, dy, \quad x \in \bd\Omega.
\end{align}
Once the boundary data of $v$ is computed, we have
\begin{align}
  \label{eqn:cauchy}
  v(x) = \frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{v(y)}{y-x} \,dy, \quad
  v'(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^2} \, dy, \quad
  v''(x) = \frac{1}{\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^3} \, dy.
\end{align}
for $x \in \Omega$.  Since $v(x)$ depends on the density function $\eta
\in \CC$, we use the notation $v[\eta](x)$ to denote the holomorphic
function defined in equation~\eqref{eqn:laplaceComplex}, and the
derivatives are written as $v'[\eta](x)$ and $v''[\eta](x)$.

The Stokes double-layer potential~\eqref{eqn:stokesDLP} can be
decomposed as
All these integrals but the first are a gradient of a Laplace layer
potential.  In addition, each component of the first integral can be
written as a Cauchy integral with a complex-valued density
function~\cite{bar-wu-vee2015}.  Therefore, the Stokes double-layer
potential can be written as five Cauchy integrals and their derivatives
with the density functions $\yy \cdot \eeta$, $\eta_1$, and $\eta_2$.
The individual velocity components are
\begin{align}
  \label{eqn:cauchy1}
  u_1(x) &= \Re (v[\tau_1](x)) + \Re (v'[y\cdot\eta](x)) 
           -x_1\Re (v'[\eta_1](x)) - x_2\Re (v'[\eta_2](x)), \\
  u_2(x) &= \Re (v[\tau_2](x)) - \Im (v'[y\cdot\eta](x)) 
       +x_1\Im (v'[\eta_1](x)) + x_2\Im (v'[\eta_2](x)),
\end{align}
where $y \cdot \eta = y_1 \eta_1 + i y_2 \eta_2 \in \CC$, 
\begin{align} 
  \tau_1=(\eta_1+i\eta_2)\frac{\Re(n)}{n}, \quad \text{and} \quad
  \tau_2=(\eta_1+i\eta_2)\frac{\Im(n)}{n},
\end{align}
and $n \in \CC$ is the complex representation of the outward unit normal
$\nn \in \RR^2$

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Shear Stress}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
To compute the rate of erosion, we must compute the shear stress of the
double-layer potential
\begin{align}
  \tau(\xx) = \left(\nabla \DDD[\eeta](\xx) + 
    \nabla^T \DDD[\eeta](\xx) \right) \nn \cdot \ss,
\end{align} where $\nn$ and $\ss$ are the unit normal and tangent
vectors at $\xx$, respectively.  Each term of the gradient of the
velocity field can be written in terms of Cauchy integrals.  For
example,
\begin{align}
  \pderiv{u_1}{x_2} &= \Re(iv'[\tau_1](x)) + 
      \Re(i v''[y \cdot \eta](x)) -
      x_1\Re(i v''[\eta_1](x))  -
      \Re(v'[\eta_2](x)) - x_2 \Re(iv''[\eta_2](x)) \\
    &= -\Im(v''[\tau_1](x)) - 
      \Im(v''[y \cdot \eta](x)) +
      x_1\Im(v''[\eta_1](x)) -
      \Re(v'[\eta_2](x)) + x_2 \Im(v''[\eta_2](x)).
\end{align}
Similarly, the other three components of $\nabla \DDD[\eeta]$ can be
calculated
\begin{align}
  \pderiv{u_1}{x_1} &= +\Re (v'[\tau_1](x)) + 
    \Re (v''[y\cdot\eta](x)) - \Re (v'[\eta_1](x)) - 
    x_1\Re (v''[\eta_1](x)) - x_2\Re (v''[\eta_2](x)), \\
  \pderiv{u_1}{x_2} &= - \Im (v'[\tau_1](x)) - 
    \Im (v''[y\cdot\eta](x)) + x_1\Im (v''[\eta_1](x)) - 
    \Re (v'[\eta_2](x)) + x_2\Im (v''[\eta_2](x)), \\
  \pderiv{u_2}{x_1} &= +\Re (v'[\tau_2](x)) - 
    \Im (v''[y\cdot\eta](x)) + \Im (v'[\eta_1](x)) +
    x_1\Im (v''[\eta_1](x)) + x_2\Im (v''[\eta_2](x)), \\
  \pderiv{u_2}{x_2} &= -\Im (v'[\tau_2](x)) - 
    \Re (v''[y\cdot\eta](x)) + x_1\Re (v''[\eta_1](x)) +
    \Im (v'[\eta_2](x)) + x_2\Re (v''[\eta_2](x)).
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Vorticity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Having calculated all derivatives of the double-layer potential, the
vorticity can written in terms of Cauchy integrals.  In particular,
\begin{align}
  \omega(x) = \pderiv{u_2}{x_1} - \pderiv{u_1}{x_2} = 
\Re (v'[\tau_2](x)) + \Im (v'[\tau_1](x)) 
 + \Re (v'[\eta_2](x))+ \Im (v'[\eta_1](x)).
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Pressure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The pressure of the double-layer potential is
\begin{align}
  p(\xx) = -\frac{1}{\pi} \int_{\bd\Omega} \frac{1}{\rho^2}
    \left(1 - 2 \frac{\rr \otimes \rr}{\rho^2} \right) \nn
    \cdot \eeta(\yy) \, ds_\yy.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transport, Tracers, and Tortuosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Erosion leads to interesting porous geometries, and we are interested in
characterizing transport in these geometries.  To characterize the
transport, we first require tracer trajectories (streamlines) that are
governed by
\begin{align}
  \label{eqn:trajectories}
  \frac{d\ss}{dt} = \uu(\ss,t), \quad \ss(0) = \ss_0.
\end{align}
Instead of allowing the transport and erosion to happen simultaneously,
we freeze the geometry and compute the streamlines in this fixed
geometry.  Therefore, the velocity $\uu$ has no temporal dependence when
we characterize the transport.

Key variables that are used to characterize porous media flow is the
porosity (area of void space), permeability, and tortuosity.  The
tortuosity, which  quantifies the amount of turning made by a
streamline, is
\begin{align*}
  T = \frac{1}{d}\frac{\displaystyle\int_{S}u_x(y)\lambda(y)\,dy}
  {\displaystyle\int_{S}u_x(y)\,dy},
\end{align*}
where $S$ is an cross-section perpendicular to the $x$ axis with length
$d$, $\lambda(y)$ is the length of the streamline that is initialized
at the $y$-coordinate along $S$, and $u_x(y)$ is the $x$ component of
the velocity at the $y$-coordinate along $S$.  There have been efforts
to relate the tortuosity to the porosity~\cite{matyka2008tortuosity},
but these results can not be universal since the porosity which depends
on only a single number, the tortuosity is affected by the distribution
of pores within the porous network.  In addition to the hydraulic
tortuosity, a local tortuosity can be used to analyze the local behavior
of a streamline near a body~\cite{matyka2008tortuosity}.  The local
tortuosity of a streamline initialized at $y$ along $S$ is
\begin{align*}
  \tau(y) = \frac{\lambda(y)}{d}.
\end{align*}

>>>>

Since erosion alters the geometry, it also affects coarse-grained
variables.  In this paper, we examine two particular variables that can
be used to understand transport in porous media---the tortuosity and the
anomalous diffusion rates.   The tortuosity quantifies the twisting of
particle trajectories in a medium, and in laminar flow, the tortuosity
is one.  While a porous media often has a large tortuosity, as erosion
increases the gaps between bodies, preferential channels are created,
large numbers of particles become more laminar, and we expect the
tortuosity to decrease.  Anomalous diffusion is often used to describe
transport in a porous media, and it quantifies the dispersion of
particles in the flow.  Again, as the geometry erodes, particles behave
more as a ballistic particle, and we expect less dispersion.  In this
paper, we compute both the tortuosity and the anomalous diffusion rates
of several geometries that form via erosion.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Boundary Integral Equation Formulation}
\label{s:formulation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We start by defining the main variables used to model erosion.  We
consider flows inside a confined geometry $\Omega$ that contains $M$
eroding bodies with boundary $\gamma_\ell$.  The boundary of the
bounding geometry is $\Gamma$, and the fluid domain is bounded by $\bd
\Omega = \Gamma \cup \gamma_1 \cup \cdots \cup \gamma_M$.  On each
eroding body, $\nn$ is the normal vector pointing into the body, and
$\ss$ is the unit tangent vector pointing in the counterclockwise
direction.  The normal vector on $\Gamma$ is the outward unit normal.
Neglecting inertial forces, the governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{\em conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{\em conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{\em no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{\em outer wall
velocity} \\
\Vn = \CE \, \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
&&\mbox{\em erosion model}.
\end{split}
\end{equation}
Here, $\uu$ is the fluid velocity, $p$ is the pressure, $\UU$ is a
prescribed Hagen-Poiseuille velocity field, and
\begin{align}
  \tau = -(\nabla \uu + \nabla \uu^T) \nn \cdot \ss
  \label{eqn:shearStress}
\end{align}
is the shear stress on the eroding bodies.  Using the same approach as
our previous work~\cite{qua-moo2018}, we use a double-layer potential
formulation for the velocity $\uu$.  The double-layer potential is
\begin{align}
  \uu(\xx) = \DDD[\eeta](\xx) = 
    \int_{\bd\Omega} D(\xx,\yy) \eeta(\yy)\, ds_\yy = 
  \frac{1}{\pi}\int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
  \label{eqn:stokesDLP}
\end{align}
where $D$ is the kernel of the integral operator, $\rr = \xx - \yy$,
$\rho = \|\rr\|$, $\nn$ is the unit outward normal at $\yy$, and $\eeta$
is an unknown density function.  Since the Stokes double-layer potential
can not represent all solutions of the incompressible Stokes equations,
we complete the integral equation formulation by adding the Stokeslets
and rotlets~\cite{pow-mir1987}
\begin{align}
  S[\llambda_\ell,\cc_\ell](\xx) = \frac{1}{4\pi} \left( 
    -\log \rho \II + \frac{\rr \otimes \rr}{\rho^2} \right), \quad
  R[\xi_\ell,\cc_\ell](\xx) = \frac{\rr^\perp}{\rho^2} \xi_\ell,
\end{align}
where $\rr = \xx - \cc_\ell$, $\cc_\ell$ is a point inside the
$\ell^{th}$ eroding body, $\rr^\perp = (r_2,-r_1)$, and $\rho =
\|\rr\|$.  Then, the solution of the Stokes equation is
\begin{align}
  \uu(\xx) = \DDD[\eeta](\xx) + 
    \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx) + 
    \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx),
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DDD[\eeta](\xx) + 
    \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx) + 
    \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx) +
    \NN_0[\eeta](\xx), \quad \xx \in \bd\Omega, \\
  \llambda_\ell &= \frac{1}{2\pi} \int_{\gamma_\ell} 
    \eeta(\yy)\, ds_\yy \quad \ell = 1,\ldots,M, \\
  \xi_\ell &= \frac{1}{2\pi} \int_{\gamma_\ell}
    (\yy - \cc_\ell)^\perp \cdot \eeta(\yy)\, ds_\yy 
    \quad \ell = 1,\ldots,M,
\end{align}
where $\ff$ is the prescribed Dirichlet boundary condition of the
velocity on the outer wall and eroding bodies, and $\NN_0$ is the
integral operator with kernel $\nn(\xx) \otimes \nn(\yy)$, $\xx,\yy \in
\Gamma$, which removes the rank one null space resulting from the
flux-free requirement of $\ff$. Once the unknown density function is
computed, the shear stress is found by first computing the deformation
tensors of the double-layer potential, Stokeslets, and rotlets, and then
forming the shear stress as defined in equation~\eqref{eqn:shearStress}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{s:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The layer potential representations of $\uu$~\eqref{eqn:stokesDLP} and
its gradient are written in terms of contour integrals in $\RR^2$. We
will be applying a quadrature method for holomorphic functions, so we
must convert these layer potentials in $\RR^2$ to contour integrals in
$\CC$.  The first step is to write the Stokes double-layer potential as
the sum of a Laplacian double-layer potential and its gradient
\begin{equation}
  \label{eqn:Stokes2Laplace}
  \begin{aligned}
    \DDD[\eeta](\xx) &= 
      \frac{1}{2\pi} \int_{\bd\Omega} 
        \frac{\nn}{\rho^2} (\rr \cdot \eeta) \, ds_\yy + 
      \frac{1}{2\pi} \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2} (\yy \cdot \eeta) \, ds_\yy \\
      &- \frac{1}{2\pi} x_1 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_1(\yy) \, ds_\yy -
      \frac{1}{2\pi} x_2 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_2(\yy) \, ds_\yy.
  \end{aligned}
\end{equation}
Next, since the Laplacian double-layer potential can be written as the
Cauchy integral
\begin{align}
  \DD[\eta](\xx) = \Re (v(x)) \quad \text{where} \quad
  v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy,
  \label{eqn:laplaceComplex}
\end{align}
where $\xx,\yy \in \RR^2$ are identified with the points $x,y \in \CC$,
the Stokes double-layer potential can be written in terms of a complex
integral and its first derivative.  In addition, the derivative of the
Stokes double-layer potential can be written in terms of a complex
integral and its first two derivatives.

Equation~\eqref{eqn:laplaceComplex} can be converted to a Cauchy
integral by first finding boundary data of $v$ which satisfies the
Sokhotski-Plemelj jump relation
\begin{align}
  \label{eqn:SPrelation}
  v(x) = -\frac{1}{2} \eta(x) + \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x-y}\, dy, \quad x \in \bd\Omega.
\end{align}
Once the boundary data of $v$ is formed, we have
\begin{align}
  \label{eqn:cauchy}
  v(x) = \frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{v(y)}{y-x} \,dy, \quad
  v'(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^2} \, dy, \quad
  v''(x) = \frac{1}{\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^3} \, dy.
\end{align}
for $x \in \Omega$.

In equation~\eqref{eqn:cauchy}, $v(x)$ depends on the density function
$\eta \in \CC$, so we use the notation $v[\eta](x)$ to denote the
holomorphic function defined in equation~\eqref{eqn:laplaceComplex}, and
the derivatives are written as $v'[\eta](x)$ and $v''[\eta](x)$.  Having
introduced this notation, the components of the Stokes double-layer
potential are~\cite{bar-wu-vee2015}
\begin{equation}
  \begin{aligned}
    u_1(x) &= \Re (v[\tau_1](x)) + \Re (v'[y\cdot\eta](x)) 
             -x_1\Re (v'[\eta_1](x)) - x_2\Re (v'[\eta_2](x)), \\
    u_2(x) &= \Re (v[\tau_2](x)) - \Im (v'[y\cdot\eta](x)) 
         +x_1\Im (v'[\eta_1](x)) + x_2\Im (v'[\eta_2](x)),
  \end{aligned}
  \label{eqn:cauchy1}
\end{equation}
where $y \cdot \eta = y_1 \eta_1 + i y_2 \eta_2 \in \CC$, 
\begin{align} 
  \tau_1=(\eta_1+i\eta_2)\frac{\Re(n)}{n}, \quad \text{and} \quad
  \tau_2=(\eta_1+i\eta_2)\frac{\Im(n)}{n},
\end{align}
and $n \in \CC$ is the complex representation of the outward unit normal
$\nn \in \RR^2$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation for the Gradient of
the Velocity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We require complex-valued representations of both the shear stress and
the vorticity.  Both these quantities depend on the gradient of the
velocity.  Each term of the gradient of the velocity field can be
written in terms of Cauchy integrals.  Taking the derivatives of
equation~\eqref{eqn:cauchy1} with respect to $x_1$ and $x_2$, we have 
\begin{align}
  \pderiv{u_1}{x_1} &= +\Re (v'[\tau_1](x)) + 
    \Re (v''[y\cdot\eta](x)) - \Re (v'[\eta_1](x)) - 
    x_1\Re (v''[\eta_1](x)) - x_2\Re (v''[\eta_2](x)), \\
  \pderiv{u_1}{x_2} &= - \Im (v'[\tau_1](x)) - 
    \Im (v''[y\cdot\eta](x)) + x_1\Im (v''[\eta_1](x)) - 
    \Re (v'[\eta_2](x)) + x_2\Im (v''[\eta_2](x)), \\
  \pderiv{u_2}{x_1} &= +\Re (v'[\tau_2](x)) - 
    \Im (v''[y\cdot\eta](x)) + \Im (v'[\eta_1](x)) +
    x_1\Im (v''[\eta_1](x)) + x_2\Im (v''[\eta_2](x)), \\
  \pderiv{u_2}{x_2} &= -\Im (v'[\tau_2](x)) - 
    \Re (v''[y\cdot\eta](x)) + x_1\Re (v''[\eta_1](x)) +
    \Im (v'[\eta_2](x)) + x_2\Re (v''[\eta_2](x)).
\end{align}
With these four expressions, the deformation tensor of the
double-layer potential, $\ssigma^D = \frac{1}{2} (\nabla \uu + \nabla
\uu^T)$, is written in terms of Cauchy integrals.  In addition, the
vorticity can be written as
\begin{align}
  \omega(x) = \pderiv{u_2}{x_1} - \pderiv{u_1}{x_2} = 
\Re (v'[\tau_2](x)) + \Im (v'[\tau_1](x)) 
 + \Re (v'[\eta_2](x))+ \Im (v'[\eta_1](x)).
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transport, Tracers, and Tortuosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Erosion leads to interesting porous geometries, and we are interested in
characterizing transport in these geometries.  For porous media, two of
the most important variables are the porosity and the permeability.
However, additional properties of the flow are revealed by computing
the anomalous diffusion rate, the tortuosity, and the distribution the
gap sizes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Anomalous Diffusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The dynamics of fluid in porous media can be characterized in terms of
anomalous diffusion.  However, the anomalous diffusion rate depends on
the properties of the geometry.  In addition to the porosity, the
distribution of the grains affects the anomalous diffusion rates.  To
quantify this rate, we require tracer trajectories that are governed by
\begin{align}
  \label{eqn:trajectories}
  \frac{d\ss}{dt} = \uu(\ss,t), \quad \ss(0) = \ss_0.
\end{align}
Given a trajectory, we define $\lambda_j(t)$ to be the distance
travelled by the particle
\begin{align}
  \lambda_j(t) = \int_{0}^t \|\ss_j(\tau)\|\, d\tau.
\end{align}
Having formed $N_p$ trajectories, the first and second ensemble moments
are
\begin{align}
  \langle \lambda \rangle (t) = 
    \frac{1}{N_p} \sum_{j=1}^{N_p} \lambda_j(t), \quad 
    \sigma_\lambda^2(t) = \frac{1}{N_p} 
    \left[\lambda_j(t) - \langle \lambda \rangle(t) \right]^2.
\end{align}
If the particle spreading scales with time as $\sigma_\lambda^2 \sim
t^\alpha$, with $\alpha < 1$, this indicates that the particle
trajectories are sub-diffusive.  In contrast, if $\sigma_\lambda^2 \sim
t^2$, the particle trajectories are undergoing a ballistic motion and
are not affected by the grains.

For early times when the geometry has not eroded, we expect the
trajectories to be sub-diffusive.  However, as the bodies erode and
straight preferential channels are formed, we expect the dynamics to
resemble a ballistic motion.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Tortuosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The tortuosity of a geometry quantifies the average amount of turning
made by a streamline.  Mathematically, it is defined as
\begin{align}
  T = \frac{1}{d}\frac{\displaystyle\int_{S}u_x(y)\lambda(y)\,dy}
  {\displaystyle\int_{S}u_x(y)\,dy},
  \label{eqn:tortuosity1}
\end{align}
where $S$ is an cross-section perpendicular to the $x$ axis with length
$d$, $\lambda(y)$ is the length of the streamline that is initialized
at the $y$-coordinate along $S$, and $u_x(y)$ is the $x$ component of
the velocity at the $y$-coordinate along $S$.  Under the assumptions
that the flow is incompressible and not reentrant, the tortuosity can be
equivalently measured as~\cite{dud-koz-mat2011}
\begin{align}
  T = \frac{\displaystyle\int_\Omega \|\uu(\yy)\|\, d\yy}
           {\displaystyle\int_\Omega u_x(\yy)\, d\yy}.
  \label{eqn:tortuosity2}
\end{align}
The flow is reentrant if all streamlines connect the inlet to the
outlet.  Recirculation zones are possible in viscous fluids, but they
only constitute a small fraction of the total porous media.  Therefore,
the tortuosity defined in equation~\eqref{eqn:tortuosity2} is an upper
bound for the definition in equation~\eqref{eqn:tortuosity1}.  We have
used formulas for a particular geometry, and the difference is
negligible.  Since equation~\eqref{eqn:tortuosity2} does not require the
additional work of computing particle trajectories, we use this
definition for the tortuosity.
